### 2019-03-12 Paper.Rmd
monday night updates

---
title: "ECON 21320: Final Paper"
author: "Eric Milstein and Tak Hogan"
date: "March 21, 2019"
bibliography: references.bib
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tigris)
library(gplots)
library(foreign)
library(car)
library(qwraps2)
library(knitr)
```

# Abstract
[@aliprantisBlowingItKnocking2015]

# Introduction

# Literature Review

# Methods

# Data and Analysis

```{r}
# 
# # Summary statistics table
# 
# our_summary1 <-
#   list("Assessed Land Value" =
#          list("min" = ~ min(total.clean$Assessed.Land.Value),
#               "median" = ~ median(total.clean$Assessed.Land.Value),
#               "max" = ~ max(total.clean$Assessed.Land.Value),
#               "mean (sd)" = ~ qwraps2::mean_sd(total.clean$Assessed.Land.Value)),
#        "Assessed Improvement Value" =
#          list("min" = ~ min(total.clean$Assessed.Improvement.Value),
#               "median" = ~ median(total.clean$Assessed.Improvement.Value),
#               "max" = ~ max(total.clean$Assessed.Improvement.Value),
#               "mean (sd)" = ~ qwraps2::mean_sd(total.clean$Assessed.Improvement.Value))
#               "Forward Gears" =
#        list("Three" = ~ qwraps2::n_perc0(total.clean$gear == 3),
#             "Four"  = ~ qwraps2::n_perc0(total.clean$gear == 4),
#             "Five"  = ~ qwraps2::n_perc0(total.clean$gear == 5))
#        )
# 
# ### Overall
# by_415 <- summary_table(total.clean, our_summary1)
# kable(by_415)

```

# References