### 2019-03-15 analysis.Rmd
edits from weds/thursday morning

---
title: "analysis"
author: "Eric Milstein"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotrix)
```

Regression output

```{r}

# m1 <- lm(`time_vals.(1)`$t_m_1_100 ~ `time_vals.(1)`$completionyear + `time_vals.(1)`$section415dec, data = `time_vals.(1)`)
# 
# m2 <- lm(~ , data = analysis.data)
# m3 <- lm(~ , data = analysis.data)
# 
# stargazer(m1, m2, m3, type="html",
#           dep.var.labels=c("Median","Fast car (=1)"),
#           covariate.labels=c("Gross horsepower","Rear axle ratio","Four foward gears", "Five
#                              forward gears","Type of transmission (manual=1)"),
#           out="models.htm")

```

Diff-in-diff graph

```{r}

#p <- ggplot(data.frame(x=c(-5, 5), y=c(-30,30)), aes(x,y)) + labs(x = "Years Before/After Affordable Housing Construction") + labs(y #= "Log(Median Property Values Within 300 Meters)") + labs(title = "Difference-in-Differences Plot for ") + #geom_vline(xintercept=0)

plot(x=NA, type="n", ylim=c(-5, 5), xlim=c(-5, 5),
     xlab="Years Before/After Affordable Housing Construction", ylab="Log(Median Property Values Within 300m)")
# Group 1
ablineclip(a=1, b=0.2, x1= -5, x2=0, col="blue")
ablineclip(a=1, b=0.4, x1= 0, x2=5, col="blue")

# Group 2
ablineclip(a=-2, b=0.15,x1 = -5, x2=0, col="red")
abline(v = 0)

ablineclip(a=-2, b=0.2, x1= 0, x2=5, col="red")

```